---
title: "data_analysis_notebook"
output: html_document
---

## 0. Description
Three-way analysis of covariants (ANCOVA) of the effect of:
 - residual bandwidth
 - graph type
 - trend type
 -----> on error of estimation

Covariantes:
 - participant ids
 - actual slope

All variables are within-subjects.
The dependent variable is the error.

Plan:
1. open the data
2. visualisation
3. ANCOVA analysis

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(TukeyC)
library(car)
library(modelr)
library(lmtest)
library(MASS)
import::from(lmerTest, lmer)
import::from(multcomp, glht, mcp, contrMat)
import::from(broom, tidy)
import::from(psycho, analyze)
```


## 1. Open the data
1.1. if **our data**
Run to join all data files into a single one (do just once: we will not include the part into the final submition):
```{r}
p1 <- read.csv("./data/P1.csv")
p2 <- read.csv("./data/P2.csv")
p3 <- read.csv("./data/P3.csv")
p4 <- read.csv("./data/P4.csv")
p5 <- read.csv("./data/P5.csv")
p6 <- read.csv("./data/P6.csv")
p7 <- read.csv("./data/P7.csv")
p8 <- read.csv("./data/P8.csv")
# p9 <- read.csv("./data/P9.csv")
p10 <- read.csv("./data/P10.csv")
p11 <- read.csv("./data/P11.csv")
p12 <- read.csv("./data/P12.csv")
p13 <- read.csv("./data/P13.csv")
p14 <- read.csv("./data/P14.csv")
# p15 <- read.csv("./data/P15.csv")
p16 <- read.csv("./data/P16.csv")
p17 <- read.csv("./data/P17.csv")
p18 <- read.csv("./data/P18.csv")
# merge data from all participant in one df
our_res <- rbind(p1, p2, p3, p4, p5, p6, p7, p8, p10, p11, p12, p13, p14, p16, p17, p18)
# to filter from the validation stimuli
our_res <- filter(our_res, isValidation == 0)
# Write all data to a .csv file:
write.csv(our_res, file = "./data/trend-bias-16-results.csv")
```

Open a file with all our results:
```{r}
all_results <- read.csv("../data/trend-bias-16-results.csv")
```

1.2. if the **original data** (from the paper)
```{r}
orig_all_results <- read.csv("../data/original/trend-bias-adjusted.csv")
```


Taking the **quartiles**:
```{r}
all_results <- subset(all_results, all_results$Error..unsigned. >= quantile(all_results$Error..unsigned., 0.25 ) & all_results$Error..unsigned. <= quantile(all_results$Error..unsigned. , 0.75 ))
```

> colnames(p1)
"Sigma", "Type", "M", "Participant", "Graph.Type", "Error",
"Error..unsigned.", "Index", "Filepath", "isValidation"

Enter robust linear regression with all the variables for a first impression:
```{r}
ols <- lm(Error..unsigned. ~ Sigma + Type + M + Graph.Type + Participant, data=all_results)
aov <- aov(Error..unsigned. ~ Sigma + Type + M + Graph.Type + Participant, data=all_results)
summary(ols)
# robust <- coeftest(ols, vcov=vcovHC(ols))
# robust
```


## 2. Visualize the data
smean.cl.normal(...) in Hmisc (wrapped in tidyverse), which uses quantiles of the t-distribution to calculate confidence limits.

Build three plots together (bandwidth of residuals,graph type and trend type):
```{r}
# 2.1. Bandwidth of Residuals (Sigma)
residuals_plot <- 
  all_results %>% 
   ggplot(aes(x = Sigma, y = Error..unsigned.)) +
       geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
       stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
       ylab("Mean(error)") +
       xlab("Bandwidth of Residuals") +
       expand_limits(y = 0) + coord_flip()
residuals_plot

# 2.2. Graph type (Graph.Type)
graph_plot <- 
  all_results %>% 
    ggplot(aes(x = Graph.Type, y = Error..unsigned.)) +
        geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
        stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
        ylab("Mean(error)") +
        xlab("Chart type") +
        expand_limits(y = 0) + coord_flip()
graph_plot

# 2.3. Trend type (Type)
trend_plot <- 
  all_results %>% 
    ggplot(aes(x = Type, y = Error..unsigned.)) +
        geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
        stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
        ylab("Mean(error)") +
        xlab("Trend type") +
        expand_limits(y = 0) + coord_flip()
trend_plot

grid.arrange(residuals_plot, graph_plot, trend_plot, nrow = 1)
```


## 3. Analysis
3.1. Carry out **three way ANCOVA** (lmer and for "mixed effect" (appears because of the difference between participants) is used)
```{r}
ols1 <- lmer(Error..unsigned. ~ Sigma*Graph.Type*Type + M + (1|Participant), data=all_results)
ancova <- anova(ols1)
print(ancova)  # Interaction is not significant
```


*** 3.2.-3.4. include the analysis for each of the independent variables separately,
but now I do not think that we need them. Three way ANCOVA in 3.1. should be enough.
3.2. Residual bandwidth
Regression model with "Error..unsigned." as the response variable, "Sigma" as the predictor variable, and "Participant" and "M" (=slope) as the categorical variables.
With interactions between categorical variables and predictor variable are taken into account:
```{r}
model_residual_bandwidth <- aov(Error..unsigned.~Sigma*Participant*M, data=all_results)
print(summary(model_residual_bandwidth))

# Without interactions between categorical variables and predictor variable are taken into account:
model_residual_bandwidth2 <- aov(Error..unsigned.~Sigma+Participant+M, data=all_results)
print(summary(model_residual_bandwidth2))  # Interaction is not significant

# Compare the two models:
print(anova(model_residual_bandwidth, model_residual_bandwidth2))
```

3.3. Graph type
Regression model with "Error" as the response variable, "Graph.Type" as the predictor variable, and "Participant" and "M" (=slope) as the categorical variables.
With interactions between categorical variables and predictor variable are taken into account:
```{r}
model_graph_type <- aov(Error~Graph.Type*Participant*M, data=q_results)
print(summary(model_graph_type))
# Without interactions between categorical variables and predictor variable are taken into account:
model_graph_type2 <- aov(Error~Graph.Type+Participant+M, data=q_results)
print(summary(model_graph_type2))

# Compare the two models.
print(anova(model_graph_type, model_graph_type2)) # Interaction is not significant
# plot(model_graph_type2)
```

3.4. Trend type
Regression model with "Error..unsigned." as the response variable, "Type" as the predictor variable, and "Participant" and "M" (=slope) as the categorical variables.
With interactions between categorical variables and predictor variable are taken into account:
```{r}
model_trend_type <- aov(Error..unsigned.~Type*Participant*M, data=q_results)
print(summary(model_trend_type))
# Without interactions between categorical variables and predictor variable are taken into account:
model_trend_type2 <- aov(Error..unsigned.~Type+Participant+M, data=q_results)
print(summary(model_trend_type2))

# Compare the two models.
print(anova(model_trend_type, model_trend_type2))  # Interaction is not significant

# plot(Error..unsigned.~Participant, all_results)
```


# 4. Futher analysis
4.1. Average signed error
```{r}
mean(all_results$Error)
```
One-sample t-test. H0: mean of signed error is null:
```{r}
t.test(all_results$Error)
```
This means that on our data the H1 is true...

4.2. **Tukey's Honest Significant Difference** for Trend types
```{r}
TukeyHSD(aov, "Type", ordered = TRUE)
plot(TukeyHSD(aov, "Type"))  # no difference in accuracy
```

4.3. **dropterm()**
```{r}
dropterm(ols1, test='Chisq')
```
