Team H
Authors: Mervin Cheok, Elfat Esati, Jan Gugler, Thomas Hugentobler, Lorenz Nagele, Iuliia Nigmatulina

---
title: "Regression by Eye: Data Analysis"
output: html_document
---

## 0. Description

Three-way analysis of covariance (ANCOVA) of the effect of:
 - residual bandwidth
 - graph type
 - trend type
 
Measure: absolute error of estimation

Covariantes:
 - participant ids
 - actual slope

All variables are within-subjects design.
Dependent variable: error

Inference plan:
1. Load and clean the data
2. Visualisations
3. ANCOVA analysis

```{r}
library(tidyverse)
library(ggplot2)
library(grid)
library(gridExtra)
library(TukeyC)
library(car)
library(modelr)
library(lmtest)
library(MASS)
import::from(lmerTest, lmer)
import::from(multcomp, glht, mcp, contrMat)
import::from(broom, tidy)
import::from(psycho, analyze)
```


## 1. Loading and pre-processing the data

1.1. **Our data**
Run to join all data files into a single one (do just once: we will not include the part into the final submission):
```{r}
p1 <- read.csv("../data/P1.csv")
p2 <- read.csv("../data/P2.csv")
p3 <- read.csv("../data/P3.csv")
p4 <- read.csv("../data/P4.csv")
p5 <- read.csv("../data/P5.csv")
p6 <- read.csv("../data/P6.csv")
p7 <- read.csv("../data/P7.csv")
p8 <- read.csv("../data/P8.csv")
# p9 <- read.csv("../data/P9.csv")
p10 <- read.csv("../data/P10.csv")
p11 <- read.csv("../data/P11.csv")
p12 <- read.csv("../data/P12.csv")
p13 <- read.csv("../data/P13.csv")
p14 <- read.csv("../data/P14.csv")
# p15 <- read.csv("../data/P15.csv")
p16 <- read.csv("../data/P16.csv")
p17 <- read.csv("../data/P17.csv")
p18 <- read.csv("../data/P18.csv")

# merge data from all participant in one df
all_res <- rbind(p1, p2, p3, p4, p5, p6, p7, p8, p10, p11, p12, p13, p14, p16, p17, p18)
rm(p1, p2, p3, p4, p5, p6, p7, p8, p10, p11, p12, p13, p14, p16, p17, p18)
```

```{r}
# check the validation stimuli and delete those participants whose average error exceeds 0.2
validation_stimuli <- all_res %>% filter(isValidation == 1) %>% group_by(Participant) %>% summarise(mean = mean(Error..unsigned.))
validation_stimuli_passed <- filter(validation_stimuli, mean < 0.2)
filtered_res <- all_res %>% filter(isValidation == 0) %>% inner_join(validation_stimuli_passed, by="Participant")
```

```{r}
# Write data to a .csv file:
write.csv(filtered_res, file = "../data/trend-bias-results.csv")
rm(validation_stimuli, validation_stimuli_passed, all_res, filtered_res)
```

Open a file with complete results:
```{r}
our_results <- read.csv("../data/trend-bias-results.csv")
```

1.2. **Original data** (from the paper)
```{r}
orig_results <- read.csv("../data/trend-bias-adjusted.csv")
```


Taking the **quartiles**:
```{r}
quartiled_results <- subset(our_results, our_results$Error..unsigned. >= quantile(our_results$Error..unsigned., 0.25 ) & our_results$Error..unsigned. <= quantile(our_results$Error..unsigned. , 0.75 ))
```

```{r}
orig_quartiled_results <- subset(orig_results, orig_results$Error..unsigned. >= quantile(orig_results$Error..unsigned., 0.25 ) & orig_results$Error..unsigned. <= quantile(orig_results$Error..unsigned. , 0.75 ))
```

> colnames(p1)
"Sigma", "Type", "M", "Participant", "Graph.Type", "Error",
"Error..unsigned.", "Index", "Filepath", "isValidation"

Enter robust linear regression with all the variables for a first impression:
```{r}
ols <- lm(Error..unsigned. ~ Sigma + Type + M + Graph.Type + Participant, data=our_results)
aov <- aov(Error..unsigned. ~ Sigma + Type + M + Graph.Type + Participant, data=our_results)
summary(ols)
# robust <- coeftest(ols, vcov=vcovHC(ols))
# robust
```


## 2. Visualise the data

Create a **function** to plot three graphs (bandwidth of residuals, graph type and trend type) together:
```{r}
# collect all graphs in one variable
plot_graphs <- function(x, title) {
  
      # 2.1. Bandwidth of Residuals (Sigma)
      residuals_plot <- 
        x %>% 
         ggplot(aes(x = Sigma, y = Error..unsigned.)) +
             geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
             stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
             ylab("Mean(error)") +
             xlab("Bandwidth of Residuals") +
             expand_limits(y = 0) + coord_flip()
      # residuals_plot
      
      # 2.2. Graph type (Graph.Type)
      graph_plot <- 
        x %>% 
          ggplot(aes(x = Graph.Type, y = Error..unsigned.)) +
              geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
              stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
              ylab("Mean(error)") +
              xlab("Chart type") +
              expand_limits(y = 0) + coord_flip()
      # graph_plot
      
      # 2.3. Trend type (Type)
      trend_plot <- 
        x %>% 
          ggplot(aes(x = Type, y = Error..unsigned.)) +
              geom_point(stat = "summary", fun.y = "mean", size = 2, color = "red") +
              stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0) + 
              ylab("Mean(error)") +
              xlab("Trend type") +
              expand_limits(y = 0) + coord_flip()
      # trend_plot
      
      grid.arrange(residuals_plot, graph_plot, trend_plot, nrow = 1, top=textGrob(title, gp=gpar(fontsize=15, font=8)))
}
```

Visualise the original data:
```{r}
plot_graphs(orig_results, "Visualisation of the original data")
```

Visualise our data:
```{r}
plot_graphs(our_results, "Visualisation of our data")
```


## 3. Analysis
3.1. Carry out **three way ANCOVA** (lmer and for "mixed-effect model" (appears because of the difference between participants) is used):
- the original data
```{r}
ols_orig <- lmer(Error..unsigned. ~ Sigma*Graph.Type*Type + M + (1|Participant), data=orig_quartiled_results)
ancova_orig <- anova(ols_orig)
print(ancova_orig)  # Interaction is not significant
```

- our results
```{r}
ols_our <- lmer(Error..unsigned. ~ Sigma*Graph.Type*Type + M + (1|Participant), data=quartiled_results)
ancova_our <- anova(ols_our)
print(ancova_our)  # Interaction is not significant
```


# 4. Further analysis
4.1. Average signed error
- the original data
```{r}
mean(orig_results$Error, na.rm = TRUE)
# one-sample t-test. H0: mean of signed error is null:
t.test(orig_results$Error, na.rm = TRUE)
```

- our data
```{r}
mean(our_results$Error, na.rm = TRUE)
# one-sample t-test. H0: mean of signed error is null:
t.test(our_results$Error, na.rm = TRUE)
```
This means that on our data the H1 is true...

4.2. **Tukey's Honest Significant Difference** for Trend types
```{r}
TukeyHSD(aov, "Type", ordered = TRUE)
plot(TukeyHSD(aov, "Type"))  # no difference in accuracy
```

4.3. The interquartile mean of the absolute log error across all conditions in this experiment:
- the original data **(SOMETHING IS WRONG HERE, THE RESULT IS VERY DIFFERENT!)**
```{r}
mean(log(orig_quartiled_results$Error..unsigned.))
```
- our data
```{r}
mean(log(quartiled_results$Error..unsigned.))
```